Compositional equivalences based on Open pNets

### 2022-04-07 2016-Weak-FH-Bisim/Coverletter.tex
cover letter and proof structure

\documentclass[10pt]{article}
\usepackage[latin1]{inputenc} 
\usepackage{alltt,url}
%% Old ttbox definition without frame:
\newenvironment{ttbox}{\begin{alltt}\small\tt}%
                      {\end{alltt}}
\usepackage{color}
\newcommand{\Eric}[1]{\textcolor{blue}{#1}}
\thispagestyle{empty}
\addtolength{\textheight}{2cm}
\begin{document}
\section*{Cover Letter }
\subsection*{Submission to Logical Methods in Computer Science}
\medskip
\begin{ttbox}
Title: Compositional equivalences based on Open pNets
Authors: R. Ameur-Boulifa, L. Henrio, and E. Madelaine
\end{ttbox}
\bigskip
%\subsubsection*{Letter to Reviewers and Editors}
Dear reviewers and editors,

\medskip
This article defines bisimulation relations  for the comparison of systems specified as parameterised Networks of synchronised automata; it formalises the work we did in the last 4 years and provides a theory for the operators composing concurrent processes. This theory is based on the definition of operators as open processes. We define the operational semantics of open processes, using open transitions that include symbolic hypotheses on the behaviour of the holes. The  semantics of pNets is given by means of a symbolic open transition system. Behavioural equivalences for pNets, like strong and weak bisimilarities, are defined over such a open transition systems. We finally prove that these equivalences are c
ompositional.

@@

We think this article fits with the scope and criteria of ``Logical Methods in Computer Science'' for the following reasons. First, to our knowledge this is the first paper proposing a theoretical framework for analysing the behaviour of open systems modeled as pNets, i.e., open terms of suitable process calculi.
Second, the quite theoretical and fundamental approach adopted in this paper fits well to the topics of LMCS.
Finally, because we already applied pNets and their equivalence relations to the verification of concrete cases and because we illustrate our approach on a small but realistic example, we believe our article illustrates well the applicability of the theory we propose.

@@

Considering the length of the submitted material,
 we   suggest to
keep the two first appendices in the final version of the paper (leading to a
paper of 66 pages)  because they are important to understand the paper and contain the proofs.
The other appendix  will be available as a research report, we keep
them in the submission to ease the work of the reviewers. 
Of course, depending on the opinion of the reviewers, we could also move more material to  the body of the paper, or remove more proofs from the published version and only keep them in the report for example.

Finally, after a brief exchange with Alexandra Silva, we improved the form of the paper as follows. We improved the spacings at some places (e.g. between superscripts and rest of text) and tracked overfull boxes. We also adopted a reduced number of symbols and tried to reduce the use of $<....>$ to the synchronisation vectors where this notation is usual now.

>>>>

\documentclass[10pt]{article}
\usepackage[latin1]{inputenc} 
\usepackage{alltt,url}
%% Old ttbox definition without frame:
\newenvironment{ttbox}{\begin{alltt}\small\tt}%
                      {\end{alltt}}
\usepackage{color}
\newcommand{\Eric}[1]{\textcolor{blue}{#1}}
\thispagestyle{empty}
\addtolength{\textheight}{2cm}
\begin{document}
\section*{Cover Letter }
\subsection*{Submission to Journal of Logical and Algebraic Methods in Programming}
\medskip
\begin{ttbox}
Title: Compositional equivalences based on Open pNets
Authors: R. Ameur-Boulifa, L. Henrio, and E. Madelaine
\end{ttbox}
\bigskip
%\subsubsection*{Letter to Reviewers and Editors}
Dear reviewers and editors,

\medskip
This article defines bisimulation relations  for the comparison of systems specified as parameterised Networks of synchronised automata (called pNets); it formalises the work we did in the last 4 years and provides a theory for the operators composing concurrent processes. This theory is based on the definition of operators as open processes. We define the operational semantics of open processes, using open transitions that include symbolic hypotheses on the behaviour of the holes. The  semantics of pNets is given by means of a symbolic open transition system. Behavioural equivalences for pNets, like strong and weak bisimilarities, are defined over such a open transition systems. We finally prove that these equivalences are compositional.

@@

We think this article fits with the scope and criteria of ``Journal of Logical and Algebraic Methods in Programming'' for the following reasons. First, to our knowledge this is the first paper proposing a theoretical framework for analysing the behaviour of open systems modeled as pNets, i.e., open terms of suitable process calculi.
Second, the quite theoretical and fundamental approach adopted in this paper fits well to the topics of JLAMP.
Also, we believe our article illustrates well the applicability of the theory we propose.
Finally, the semantics of closed pNets and their usage was published in JLAMP in 2017; this submission extends the previous work to open systems and consequently provides a more symbolic semantics.

@@

Considering the length of the submitted material, the article is 40 pages long but with 50 pages of appendices. The first 2 appendices (approx. 35 pages) are detailed proofs, and the last one (approx 15 pages) is details on the usecase.
Depending on the policy of the journal concerning paper length and on the opinion of the reviewers we could either publish only the body of the paper, or keep the two first appendices, or keep the 90 pages.
In any case, all appendices will appear in a research report.

### 2022-04-07 2016-Weak-FH-Bisim/WeakBisim.tex
cover letter and proof structure

Concerning predicates, we also separate predicates on $P$ from predicates on $Q$, and from the equality on the action filling the hole:
%\begin{small}
%\begin{equation*}
%\begin{split}
%\Pred\,''&=\bigwedge_{i=0}^{m_1}\Pred\,'_{1i}\subst{\bigodot_{j=i-1}^{0}\Post\,'_{1j}}\land\Pred\,'_2 \subst{\bigodot_{j=m_1}^{0}\Post\,'_{1j}}\land\\
%&~~{\bigwedge_{i=0}^{m_3}\Pred\,'_{3i}\subst{\bigodot_{j=i-1}^{0}\Post\,'_{3j}\shortotimes\Post\,'_2\shortotimes\bigodot_{j=m_1}^{0}\Post\,'_{1j}}}\\
%&=\big(\bigwedge_{p=0}^{m_1}\!\Pred_{1p}\subst{\bigodot_{j=p-1}^{0}\Post_{1j}}\land\Pred_2\subst{\bigodot_{p=m_1}^{0}\Post_{1p}}\land\\
%&~~~\bigwedge_{p=0}^{m_2}\!\Pred_{3p}\subst{\bigodot_{j=p-1}^{0}\Post_{3j}\otimes\Post_2 \otimes\bigodot_{p=m_1}^{0}\Post_{1p}}\big) \land \bigwedge_{t=0}^{n} \Pred_t\bigodot_{i=t-1}^{0}\Post_i \land
%%$~\qquad\qquad\,\displaystyle{\bigwedge_{i=0}^{m-1}(\beta_{1{n_i}{j_0}}=\alpha_i)\subst{\bigodot_{j={n_i}-1}^{0}\Post\,'_{1j}}\land(\beta_{2{j_0}}=\alpha_m) \subst{\bigodot_{j=m_1}^{0}\Post\,'_{1j}}\land}
%\end{split}
%\end{equation*}
%%$~\qquad\qquad\,\displaystyle{\bigwedge_{i=m+1}^{n}(\beta_{3{n'_i}{j_0}}=\alpha_i)\subst{\bigodot_{j=n'_i-1}^{0}\Post\,'_{3j}\shortotimes\Post\,'_2\shortotimes\bigodot_{j=m_1}^{0}\Post\,'_{1j}}\Big)}$\\
%\end{small}
\begin{footnotesize}
\begin{align*}
\Pred\,''&=\Big(\bigwedge_{i=0}^{m_1}\Pred\,'_{1i}\subst{\bigodot_{j=i-1}^{0}\!\!\Post\,'_{1j}}\land\Pred\,'_2 \subst{\bigodot_{j=m_1}^{0}\!\!\Post\,'_{1j}}\land%\\ &\qquad
%~\quad
\bigwedge_{i=0}^{m_3}\Pred\,'_{3i}\subst{\bigodot_{j=i-1}^{0}\!\!\Post\,'_{3j}\shortodot\Post\,'_2\shortodot\bigodot_{j=m_1}^{0}\!\Post\,'_{1j}}\Big)\\
&=\Big(\bigwedge_{p=0}^{m_1}\!\Pred_{1p}\subst{\!\!\bigodot_{j=p-1}^{0}\!\!\Post_{1j}}\land\Pred_2\subst{\!\!\bigodot_{p=m_1}^{0}\!\!\Post_{1p}}\land \bigwedge_{p=0}^{m_2}\!\Pred_{3p}\subst{\!\!\bigodot_{j=p-1}^{0}\!\!\Post_{3j}\otimes\Post_2 \otimes\!\!\bigodot_{p=m_1}^{0}\!\!\Post_{1p}}\Big)\\&\qquad
\land \bigwedge_{t=0}^{n} \Pred_t\bigodot_{i=t-1}^{0}\!\!\Post_i%\\&\qquad
\land \Big( \bigwedge_{i=0}^{m-1}(\beta_{1{n_i}{j_0}}=\alpha_i)\subst{\bigodot_{j={n_i}-1}^{0}\Post\,'_{1j}}\land(\beta_{2{j_0}}=\alpha_m) \subst{\bigodot_{j=m_1}^{0}\!\!\Post\,'_{1j}}\land\\ &\qquad
~\quad\bigwedge_{i=m+1}^{n}(\beta_{3{n'_i}{j_0}}=\alpha_i)\subst{\bigodot_{j=n'_i-1}^{0}\Post\,'_{3j}\shortodot\Post\,'_2\shortodot\bigodot_{j=m_1}^{0}\!\!\Post\,'_{1j}}\Big)\displaybreak[0]\\
&=\Big(\bigwedge_{p=0}^{m_1}\!\Pred_{1p}\subst{\!\!\bigodot_{j=p-1}^{0}\!\!\Post_{1j}}\big)\land\Pred_2\subst{\!\!\bigodot_{p=m_1}^{0}\!\!\Post_{1p}}\land \bigwedge_{p=0}^{m_2}\!\Pred_{3p}\subst{\!\!\bigodot_{j=p-1}^{0}\!\!\Post_{3j}\otimes\Post_2 \otimes\!\!\bigodot_{p=m_1}^{0}\!\!\Post_{1p}}\Big)\\&\qquad 
\land \bigwedge_{t=0}^{n} \Pred_t\bigodot_{i=t-1}^{0}\!\!\Post_i
\land \Big(\bigwedge_{i=0}^{m-1}(\beta_{i}=\alpha_i)\subst{\bigodot_{j={i-1}}^{0}\!\!\Post_{j}}\land(\beta_m=\alpha_m )\subst{\bigodot_{j=m}^{0}\Post_{j}}\land\\ &\qquad
~\quad\bigwedge_{i=m+1}^{n}(\beta_i=\alpha_i)\subst{\bigodot_{j=i-1}^{m}\!\!\Post_j\shortodot\Post_m\shortodot\bigodot_{j=m-1}^{0}\!\!\Post_{j}}\Big)\\
&=\Pred
\end{align*}
\end{footnotesize}

@@

\bigskip
Consider the first triple <b0, 000, True>, we have to prove the following 6 properties, in which $OT<<WOT$ means that the (strong) open transition $OT$ is covered, in the sense
of definition \ref{def-Weak-bisim}, by the weak transition $WOT$ (it could be a set, but this will not be used here):

>>>>

Concerning predicates, we also separate predicates on $P$ from predicates on $Q$, and from the equality on the action filling the hole:
%\begin{small}
%\begin{equation*}
%\begin{split}
%\Pred\,''&=\bigwedge_{i=0}^{m_1}\Pred\,'_{1i}\subst{\bigodot_{j=i-1}^{0}\Post\,'_{1j}}\land\Pred\,'_2 \subst{\bigodot_{j=m_1}^{0}\Post\,'_{1j}}\land\\
%&~~{\bigwedge_{i=0}^{m_3}\Pred\,'_{3i}\subst{\bigodot_{j=i-1}^{0}\Post\,'_{3j}\shortotimes\Post\,'_2\shortotimes\bigodot_{j=m_1}^{0}\Post\,'_{1j}}}\\
%&=\big(\bigwedge_{p=0}^{m_1}\!\Pred_{1p}\subst{\bigodot_{j=p-1}^{0}\Post_{1j}}\land\Pred_2\subst{\bigodot_{p=m_1}^{0}\Post_{1p}}\land\\
%&~~~\bigwedge_{p=0}^{m_2}\!\Pred_{3p}\subst{\bigodot_{j=p-1}^{0}\Post_{3j}\otimes\Post_2 \otimes\bigodot_{p=m_1}^{0}\Post_{1p}}\big) \land \bigwedge_{t=0}^{n} \Pred_t\bigodot_{i=t-1}^{0}\Post_i \land
%%$~\qquad\qquad\,\displaystyle{\bigwedge_{i=0}^{m-1}(\beta_{1{n_i}{j_0}}=\alpha_i)\subst{\bigodot_{j={n_i}-1}^{0}\Post\,'_{1j}}\land(\beta_{2{j_0}}=\alpha_m) \subst{\bigodot_{j=m_1}^{0}\Post\,'_{1j}}\land}
%\end{split}
%\end{equation*}
%%$~\qquad\qquad\,\displaystyle{\bigwedge_{i=m+1}^{n}(\beta_{3{n'_i}{j_0}}=\alpha_i)\subst{\bigodot_{j=n'_i-1}^{0}\Post\,'_{3j}\shortotimes\Post\,'_2\shortotimes\bigodot_{j=m_1}^{0}\Post\,'_{1j}}\Big)}$\\
%\end{small}
\begin{footnotesize}
\begin{align*}
\Pred\,''&=\Big(\bigwedge_{i=0}^{m_1}\Pred\,'_{1i}\subst{\bigodot_{j=i-1}^{0}\!\!\Post\,'_{1j}}\land\Pred\,'_2 \subst{\bigodot_{j=m_1}^{0}\!\!\Post\,'_{1j}}\\ &\qquad\qquad \land
%~\quad
\bigwedge_{i=0}^{m_3}\Pred\,'_{3i}\subst{\bigodot_{j=i-1}^{0}\!\!\Post\,'_{3j}\shortodot\Post\,'_2\shortodot\bigodot_{j=m_1}^{0}\!\Post\,'_{1j}}\Big)\\
&=\Big(\bigwedge_{p=0}^{m_1}\!\Pred_{1p}\subst{\!\!\bigodot_{j=p-1}^{0}\!\!\Post_{1j}}\land\Pred_2\subst{\!\!\bigodot_{p=m_1}^{0}\!\!\Post_{1p}}\land \bigwedge_{p=0}^{m_2}\!\Pred_{3p}\subst{\!\!\bigodot_{j=p-1}^{0}\!\!\Post_{3j}\otimes\Post_2 \otimes\!\!\bigodot_{p=m_1}^{0}\!\!\Post_{1p}}\Big)\\&\qquad
\land \bigwedge_{t=0}^{n} \Pred_t\bigodot_{i=t-1}^{0}\!\!\Post_i%\\&\qquad
\land \Big( \bigwedge_{i=0}^{m-1}(\beta_{1{n_i}{j_0}}=\alpha_i)\subst{\bigodot_{j={n_i}-1}^{0}\Post\,'_{1j}}\land(\beta_{2{j_0}}=\alpha_m) \subst{\bigodot_{j=m_1}^{0}\!\!\Post\,'_{1j}}\\ &\qquad
~\quad\land\bigwedge_{i=m+1}^{n}(\beta_{3{n'_i}{j_0}}=\alpha_i)\subst{\bigodot_{j=n'_i-1}^{0}\Post\,'_{3j}\shortodot\Post\,'_2\shortodot\bigodot_{j=m_1}^{0}\!\!\Post\,'_{1j}}\Big)\displaybreak[0]\\
&=\Big(\bigwedge_{p=0}^{m_1}\!\Pred_{1p}\subst{\!\!\bigodot_{j=p-1}^{0}\!\!\Post_{1j}}\big)\land\Pred_2\subst{\!\!\bigodot_{p=m_1}^{0}\!\!\Post_{1p}}\land \bigwedge_{p=0}^{m_2}\!\Pred_{3p}\subst{\!\!\bigodot_{j=p-1}^{0}\!\!\Post_{3j}\otimes\Post_2 \otimes\!\!\bigodot_{p=m_1}^{0}\!\!\Post_{1p}}\Big)\\&\qquad 
\land \bigwedge_{t=0}^{n} \Pred_t\bigodot_{i=t-1}^{0}\!\!\Post_i
\land \Big(\bigwedge_{i=0}^{m-1}(\beta_{i}=\alpha_i)\subst{\bigodot_{j={i-1}}^{0}\!\!\Post_{j}}\land(\beta_m=\alpha_m )\subst{\bigodot_{j=m}^{0}\Post_{j}}\\ &\qquad
~\quad\land\bigwedge_{i=m+1}^{n}(\beta_i=\alpha_i)\subst{\bigodot_{j=i-1}^{m}\!\!\Post_j\shortodot\Post_m\shortodot\bigodot_{j=m-1}^{0}\!\!\Post_{j}}\Big)\\
&=\Pred
\end{align*}
\end{footnotesize}

@@

\bigskip
Consider the first triple <b0, 000, True>, we have to prove the following 6 properties, in which $OT<<WOT$ means that the (strong) open transition $OT$ is covered, in the sense
of Definition \ref{def-Weak-bisim}, by the weak transition $WOT$ (it could be a set, but this will not be used here):
